PANCHANGE_data_preprocessing

### 2020-08-07 PANCHANGE_data_preprocessing2020-07-15.Rmd
added logic chunk to specify cohort
got rid of .uncleaned at the demographics object, it was double

@@

# Select variables
```{r glad dem cleaning}
glad.dem.raw.id.unclean <- glad.dem.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.glad = startDate) %>% # Rename startDate
  select(
         Sample, # Sample
         ID, # ID
         startDate.glad, #Date when participant signed up to glad
         Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
         Gender = demographics.which_gender_do_you_identify_with, # Gender
         Sex = demographics.select_questionnaire_items_medical, #Sex
         EduYrs = demographics.school_education_school_years.txt, # Years in education
         demographics.college_or_university_degree, # level of education - univeristy
         demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
         demographics.o_levelsgcses_or_equivalent, # level of education - gcse
         demographics.cses_or_equivalent, # level of education - cse
         demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
         ) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.dem.raw.id.unclean)
# Inspect colnames
colnames(glad.dem.raw.id.unclean)
#Differences
dim(glad.dem.raw)[1]-dim(glad.dem.raw.id.unclean)[1]
```

@@

glad.dem.raw.id.unclean.numeric <- add_numeric(glad.dem.raw.id.unclean, exclude = exclude_cols)

# Check if factor columns have been added
colnames(glad.dem.raw.id.unclean.numeric)

# Dimensions before adding factor columns
dim(glad.dem.raw.id.unclean)
# Dimensions after adding factor columns
dim(glad.dem.raw.id.unclean.numeric)
```

@@

# identify number of outliers
length(
  which(
  glad.dem.raw.id.unclean$Age_uncleaned > age_upper_limit |
    glad.dem.raw.id.unclean$Age_uncleaned < age_lower_limit
     )
)

# remove age outliers
glad.dem.raw.id <- glad.dem.raw.id.unclean %>%
  mutate(
    Age = 
      if_else(
        Age_uncleaned > age_upper_limit | 
        Age_uncleaned < age_lower_limit, 
        true = NA_real_,
        false = Age_uncleaned,
        missing = NA_real_
      )
  )

@@

# Categorise age into groups 
In RAMP, age is indicated categorically: 16-18, 19-25, 26-35, 36-45, 46-55, 56-65, 66-70, 71-75, 76-80, 81-85, 86-90, 91-100, 100+
Age categories in GLAD, EDGI and NBR have been created to reflect this
```{r Age groups}
#if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {

@@

#}
```

>>>>

# Specify cohort
```{r}
GLAD = TRUE
EDGI = FALSE
NBR = FALSE
RAMP = FALSE
```

@@

# Select variables
```{r glad dem cleaning}
glad.dem.raw.id <- glad.dem.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.glad = startDate) %>% # Rename startDate
  select(
         Sample, # Sample
         ID, # ID
         startDate.glad, #Date when participant signed up to glad
         Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
         Gender = demographics.which_gender_do_you_identify_with, # Gender
         Sex = demographics.select_questionnaire_items_medical, #Sex
         EduYrs = demographics.school_education_school_years.txt, # Years in education
         demographics.college_or_university_degree, # level of education - univeristy
         demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
         demographics.o_levelsgcses_or_equivalent, # level of education - gcse
         demographics.cses_or_equivalent, # level of education - cse
         demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
         ) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.dem.raw.id)
# Inspect colnames
colnames(glad.dem.raw.id)
#Differences
dim(glad.dem.raw)[1]-dim(glad.dem.raw.id)[1]
```

@@

glad.dem.raw.id.numeric <- add_numeric(glad.dem.raw.id, exclude = exclude_cols)

# Check if factor columns have been added
colnames(glad.dem.raw.id.numeric)

# Dimensions before adding factor columns
dim(glad.dem.raw.id)
# Dimensions after adding factor columns
dim(glad.dem.raw.id.numeric)
```

@@

# identify number of outliers
length(
  which(
  glad.dem.raw.id$Age_uncleaned > age_upper_limit |
    glad.dem.raw.id$Age_uncleaned < age_lower_limit
     )
)

# remove age outliers
glad.dem.raw.id <- glad.dem.raw.id %>%
  mutate(
    Age = 
      if_else(
        Age_uncleaned > age_upper_limit | 
        Age_uncleaned < age_lower_limit, 
        true = NA_real_,
        false = Age_uncleaned,
        missing = NA_real_
      )
  )

@@

# Categorise age into groups 
In RAMP, age is indicated categorically: 16-18, 19-25, 26-35, 36-45, 46-55, 56-65, 66-70, 71-75, 76-80, 81-85, 86-90, 91-100, 100+
Age categories in GLAD, EDGI and NBR have been created to reflect this
```{r Age groups}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {

@@

}
```
